%!TEX root = ../template.tex
\newcommand{\novathesis}{\emph{novathesis}}
\newcommand{\novathesisclass}{\texttt{novathesis.cls}}

\chapter{Introduction}
\label{cha:introduction}

This chapter focuses on introducing the preparation phase of this dissertation. It starts by describing the motivation and the context of the problem space do be addressed. Afterwards the objectives are set, and the expected contributions to the field of study are explained. This chapter finishes by presenting the organizational structure of the reminder of the document. This dissertation is currently being developed within a corporate context, with the support of KnowledgeBiz Consulting (\gls{KBZ})~\cite{kbz}.

\section{Motivation and Context} % (fold)
\label{sec:motivation_context}
In an era where information drives our world and big data analytics are a common interest in every academic and industrial field, sports industries like Basketball, Soccer and Baseball are using more and more data in their behalf. Huge volumes of data are produced per game and per training session, and techniques like data mining and machine learning can provide coaches and players with more insightful and accurate information and analysis~\cite{nbabigdata}.

%Nike, a popular brand in the sports industry created Nike Plus an entire line of ‘products around taking previously unquantifiable athletic data and disseminating it wirelessly into easy-to-comprehend Apple-supported apps’. It is believed that the Nike plus systems boast more than 6 million "connected athletes" using sensors in running shoes, cell phones, basketball shoes, and more to ‘track movements and quantify the data while transmitting it. http://tassm.org.tw/wp-content/uploads/2018/10/Big-Data-Sport.pdf

New discoveries on data can also help the business side of sports, like the \gls{NBA} Drafts or the football transfer market.

%Falar que “Data Analysis In Sports Is Changing The Game”, tera qui uma frase que possa dar o mote a este statement. Algo deste género: The use of analytics in the sports industry will only increase as the deluge of data increases. Thanks to wearable devices, tagged equipment and video tracking, a game can generate more than a million data points. And that's just on the field -- imagine how much data flows through the stadium and beyond. As it mines new insights from all those numbers, data analysis in sports is already producing some game-changing results.


Innovations like wearable \gls{GPS} tracking devices have added big insights in team sports. Initially, practitioners of sports were interested in answering the first obvious question of the chaotic set of player movements during a game: how far did an athlete traveled over the course of a match? \gls{GPS} could easily answer this question, but soon more questions were raised. For example:
\begin{itemize}
    \item How many high velocity efforts were achieved?
    \item How much time was spent at high speed?
    \item What high accelerations and decelerations were performed? What impact does this have on the recovery strategies for each individual athlete?
\end{itemize}

This granular movements can be measured in a laboratory, using biomechanical tools like force plates that measure ground reaction forces, but this measurement tools can’t be deployed in real life cases like games or training sessions, and a “gross” positioning system like \gls{GPS} can never measure this type of forces. 

For a complete tracking solution, we need not only to analyze the “big picture” (the movement across the field), that can be provided by \gls{GPS}, but to consider the micro or local movements, which can provide enormously relevant information for performance enhancing, fatigue analysis and recovery strategy planning. 

Countless tracking systems exist for different sports, and even for different activities in each sport, but many of them are either very expensive, resource-intensive or attached to the field. Currently, there are two widely used systems to track the position and collect metrics from the players: sensors attached to each player, which have included a \gls{GPS} unit and an inertial measuring unit (accelerometer, gyroscope and magnetometer), and an array of cameras placed around the field of the game (e.g. SportVU and recently Second Spectrum used in \gls{NBA})~\cite{secondspectrumnba}.


However, these systems have particular drawbacks: 
\begin{itemize}
    \item The \gls{GPS} signal can accurately measure the position, the speed and the displacement of a player. But it can’t measure metrics like jumps, shoots and passes or falls.
    \item When in a closed court (like most of basketball courts), there is no \gls{GPS} signal. The tracking is made using cameras around the court. 
    \item This visual system can accurately track the position of the players and gather advanced metrics. The downside of this system is the need of high processing power, and that the cameras are attached to the court. In this way, every court needs to have this expensive system installed to be able to track the players performance, something that is often out of reach for low ranked and amateur teams.
    \item Most times an operator is needed to tag the events recorded by the video systems, or the video is recorded and then the processed data is provided to the coaches and players.
    
\end{itemize}

In order to tackle the drawbacks of video and \gls{GPS} solutions, Kinexon~\cite{kinexon} developed a system using a wearable sensor that can accurately track basketball players positions indoors, using Ultra-WideBand (\gls{UWB}) technology, and claims to monitor player load and efforts measured in real-time during practices or matches, providing tactic and technic analysis~\cite{kinexonarticle}.

However, this system falls short in the field of micro-movement identification, where the analysis of shoots and passes or attacks and defenses can be a useful tool for coaches to improve the performance of the team.

With this motivation, KnowledgeBiz Consulting (\gls{KBZ}) challenged me to develop a system capable of tracking the position and to collect insights on performance metrics of a single player and of the entire team, using real-time data from inertial sensors. Apart from the technical challenges, this system should also be available for every team, in any performance level, by being low-cost and easy to include in a training or match environment. \gls{KBZ} is providing the devices that contain the inertial sensors (accelerometer, gyroscope and magnetometer), and communicate via Bluetooth.


\section{Objectives} % (fold)
\label{sec:objectives}
The proposed objective of this dissertation is to develop a system based on inertial sensors like accelerometers and gyroscopes, capable of tracking player positioning and player game metrics, in order to provide insights about basketball player and team performance in real-time.

Several sports are often practiced indoors, which prevents the use of \gls{GPS} positioning. In this cases, a tracking system using inertial sensors like accelerometers or gyroscopes can be used. In order to test this tracking system, basketball was the sport chosen. Basketball is a good sport to track player movements, because it has a high number of events occurring in small time frames when comparing to other team sports like football, the number of players per team is 5 (a small number when compared to 7 players in a handball team or 11 players in a soccer team), and it’s one of the sports that invests more in new technologies (\gls{IoT}) to track players and teams.

In order to track players, the following topics need to be studied:
    \begin{itemize}
        \item What positional tracking algorithms exist?
        \item What metrics are used to measure the performance of a basketball player?
        \item What are the sensors used to collect data from sport players?
        \item What communication protocols are used to transmit this data?
    \end{itemize}

After this analysis, a body sensor wireless network will be specified and implemented, using the most suitable sensors and communication protocols for this application.

Then, the data collected by the inertial sensors will be used and analysed in order to calculate the trajectories performed by each individual player and also to provide other metrics, such as distance traveled and number of jumps. An application will be developed, capable to visually track, analyse and display individual metrics for each player, but also providing an aggregated view regarding the whole team performance, like attacking and defending strategies in the form of heatmaps. Finally, this system (body sensor network and application) will be tested and validated.

An a priori problem that will need to be tackled is the quality of the data collected the sensors, that can present reading errors or unreliable data. This depends on various factors: the sensibility and the quality of the sensors, and the location of this sensors in the body. To help correct these errors, a filtering technique can be used. Depending to detect other activities, like dribbles or shoots, more sensors will be necessary on different body parts on the metrics to be measured, different placements of the sensors on the body will have to be tested.

Summarizing, the main goal of this dissertation is to study and analyse ways to track player positions and to measure basketball game metrics, collected from inertial sensors, and to illustrate those insights to the users (e.g.\ coaches, players) through an application, providing a central location to monitor and analyze the players and the team.

\section{Expected Contributions} % (fold)
\label{sec:contribuitions}
%As a result of this dissertation, a direct contribution to Knowledgebiz is a prototype of a Basketball team tracking system, targeted to low budget teams, like college teams.

%This prototype will include an exploratory analysis of useful basketball metrics capable of being acquired by inertial sensors. A body sensor network will also be implemented, using Bluetooth Low Energy to transmit data to a mobile application. This mobile application should be able to provide easy to understand individual and collective metrics.

%The developed system will also be the development base for the H2020 Smart4Health project, that will use the same devices to measure movement from various parts of the body. The data collection and data analysis modules will be very similar to the ones developed in this thesis.

%For the field of study, a paper in international conference proceedings is expected.

The expected contributions of this dissertation are:
\begin{enumerate}
    \item Exploratory analysis of relevant basketball metrics, obtained from inertial sensor data
    \item Implementation of a body sensor using a wireless communication protocol to transmit data
    \item Development of an application that collects data from inertial sensors, analyses it and displays the insights for the players and the team
    \item Provide the development base for the H2020 Smart4Health project, in which the data collection and data analysis modules will be very similar to the ones developed in this thesis.
\end{enumerate}


\section{Document Structure} % (fold)
\label{sec:structure}
The rest of the document is organized as follows: Chapter~\ref{cha:state_art} presents the state of the art: what are the approaches currently used by professional teams regarding player tracking and data collection, what sensors are used commonly used for this application, which communication protocols are more suitable for the goals set and the related works in the field. Chapter~\ref{cha:developed_work} will describe the developed work so far, and Chapter~\ref{cha:elaboration_proposal} proposes this thesis elaboration plan.